Is Partner Modelling Mutual?

### 2015-04-01 mutual-modeling.tex
{modeling->modelling} to follow bristish english rules

\title{Is Partner Modeling Mutual?}

@@

Deep grounding or shared meaning making requires some cognitive load. For
\citet{clark1986referring}, what is important is not the individual effort made
by the receiver of a communicative act, but the overall \emph{least collaborative
effort}.  The cost of producing a perfect utterance may be higher than the cost
of repairing the problems that may arise through misunderstandings. For
instance, subjects are less careful about adapting their utterances to their
partner when they know they can provide feedback on his/her
understanding~\citep{schober1993spatial}. We introduced the notion of
\emph{optimal collaborative effort}~\citep{dillenbourg1995evolution} to stress
that misunderstanding should not be viewed as something to be avoided (if this
was possible), but as an opportunity to engage into verbalization, explanation,
negotiation, and so forth. This issue is related to the global argument
regarding cognitive load in learning activities, especially in discovery
learning environments: there is no learning without some cognitive load, but
overload may hinder learning~\citep{paas2003cognitive}. In the context of
collaborative learning, we understand the cognitive load induced by mutual
modeling as part of~\citet{schwartz1995emergence} notion of effort towards a
shared understanding. For instance, CSCL researchers expanded the use of
\emph{collaboration scripts} \fixme{REF (Aronson,?)}. A script is a pedagogical
method that collaborative learning activities in order to foster the emergence
of productive interactions such as argumentation, explanation or conflict.
Conflict-resolution scripts such as the {\sc
ArgueGraph}~\citep{dillenbourg2008mechanics} forms pair of students with
opposite opinions, which increases the difficulty of consensus building,
requiring more justifications, more negotiation, more load. Similarly, JIGSAW
scripts provide peers with different but complementary knowledge for augmenting
(reasonably) the effort group members have to engage to reach a shared solution.

@@

\section{Mutual Modeling}

In order to build a shared understanding, do partners have to build a
representation of each other's knowledge? We refer to \emph{mutual modeling} as
the process of inferring one's partner mental states. Any claim that students
carry out a detailed monitoring of their peers would be as incorrect as any
claim that they do not maintain any representation at all. If mutual modeling
had to be permanently detailed and accurate, subjects would obviously face a
huge cognitive load. Conversely, peers could not collaborate without some
minimal amount of mutual modeling. For instance, A cannot disagree with B
without knowing that B has a different opinion. The mutual model can be implicit
(A is not aware of what he knows about B), by-default (I believe that B beliefs
what I believe unless contrary evidence), opportunistic (A does not model B
unless the conversation requires it), global (A infers B's beliefs based on
categories such as age, culture or profession) and, of course, it can be
incorrect… but it can not remain empty. Dialogues include many instances of
utterances such as ``I thought he would do that'' (first level of mutual
modelling) or even ``He thought I would do that but I intended something else.''
(second level of mutual modelling).

@@

A partner model is probably not a ``box'', \ie not a monolithical representation
but rather a mosaic of information fragments about the partner, with various
grain sizes and various life cycles. This mosaic is elaborated through various
mechanisms, first for building an initial model of the partner and then for
updating this model.  As two students meet for the first time, mutual models are
initialized by the assumptions they make upon each other from cues such as
his/her membership to large categories (age, culture, profession, ...) include
stereotypes (sportsmen, junkie, business women, Swiss,...) as well as physical
appearance. Scholars studied how initial modeling impacts communication. In
their experiments on initial mutual modelling, Slugoski, Lalljee, Lamb \&
Ginsburg~\citep{slugoski1993attribution} pretended to their subjects that their
(fake) partner had or had not received the same information. They observed that
the subjects adapted their dialogue by focusing the explanation on the items
that (s)he was supposed to ignore. Brennan (1991) \fixme{ref} showed that the subjects used
different initial strategies in forming queries depending on who they were told
their partner was.

@@

After initialization, mutual models are updated along the collaborative work
through verbal and non-verbal interactions. A default inference rule is that ``my
partner agrees with me unless he disagrees'', which reject the critiques that
mutual modeling generates an unbearable cognitive load. This default rule is
superseded by the several mechanisms for monitoring and repairing one's partner
understanding: acknowledgement, continuous attention, relevance of next turns,
facial expressions including gaze signals, etc.

Mutual modeling does not occur in a vacuum but it is highly contextualized. REF
reviews the features of the collaborative situation, namely the media
(cotemporality… ), may facilitate or hamper mutual modeling. Hutchins (SILENCE)
reported a study in which a short silence between two pilots was perfectly
interpreted because it occurred in a highly constrained communication context.
Some environments are more productive than others in helping peers to detect
their misunderstandings. Roschelle~\citep{roschelle1995construction} reformulate
the design of CSCL interfaces as providing ways for peer to detect and repairs
their misunderstanding. In CSCW, researchers have explored various so-called
``awareness tools'', \ie functionalities that inform A is about B's actions that A
could not directly perceived because B was working on a different subset of the
virtual space. Different awareness tools will be addressed in this contribution
since they allowed us to manipulate experimentally the mutual modelling activity.

@@

To report experiments on mutual modeling, we suggest
the notation \model{A}{B}{X} to denote ``A knows that B knows X''. This is a
simple but useful reduction to reason on mutual modelling. This
notation does not mean A has an explicit, monolithic representation of B: it
must be understood as an abstraction referring to complex socio-cognitive
processes. Besides, we refer to the \emph{degree of accuracy} of the model as
\Model{A}{B}{X}.

@@

    \item $cost(\alpha)$ In all studies but study 4, the cost of  grounding is
        an independent variable. Study 3 uses media richness as independent
        variable, with the hypothesis that modeling emotions is ``cheaper'' with a
        richer medium, \ie  when peers can see each other.  Studies 1,2 and 4
        use awareness tools which, in principle, reduce the cost of mutual modelling, but do
        not eliminate all costs: if the tool provide A with information about
        what B does/knows, this additional information may actually increase
        cognitive load. Awareness tools constitute as a kind of mutual modelling prosthesis,
        and, like any prosthesis, they many augment mutual modelling (by facilitating it or
        even scaffolding it) or inhibit it (by making it useless).

@@

\begin{itemize}
    \item {\bf Capturing \model{A}{B}{X}}: The simplest method is to ask $A$ what (s)he
        believes about what $B$ knows, feels, intends to do, etc. This raises
        obvious methodological concerns since such a question triggers a
        modeling process beyond what it would 'naturally' be. To avoid this
        bias, one can estimate mutual modeling after task completion. Then, the
        obvious drawback are memory losses and by post-hoc reconstruction.  The
        first method was used in study 2 and the second one in the other
        studies. Another option would be to rely on external behavioural metrics
        like eye-tracking: we hypothesise for instance that the fixation time
        reflects the efforts engaged by the human to understand, hence, model,
        the others. Such an approach has however not yet been investigated in
        the presented studies.

@@

$\mathcal{H}_{2}$ is also called a second level modeling hypothesis since H
needs to monitor B to see if A understood him/her
$\mathcal{M}(B,(A,\mathcal{M}(A,B)))$

Our main issue is the symmetry questions: what is the relationship
(Fig.~\ref{mm_symmetry}) between \Model{A}{B}{X} and \Model{B}{A}{X}? A low
symmetry would mean that mutual modeling is mainly an individual
attitude/aptitude ($\mathcal{H}_{1}$). A high correlation might support
$\mathcal{H}_{2}$ since there is a low probability that randomly formed pairs
integrate peers with the same level of mutual modelling skills.

@@

This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do A and
B have the same accuracy when modeling C? $\Delta_2 = \Delta(\M{A}{C}{X},
\M{B}{C}{X})$. If it is the same, $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ gain
over $\mathcal{H}_{1}$

@@

We could go further by comparing self- versus other modeling ($\Delta_4$ in
Fig.~\ref{mm_rectangle}) as an indication of meta-cognitive skills. We can also
question if modelling skills depend upon what aspects are being modeled ($X$ or $Y$),
which would explain vertical differences ($\Delta_5$ in
Fig.~\ref{mm_rectangle}). We do not however further develop these questions because we
don't have the necessary data in our studies.

@@

This simple notation does not pretend to provide a mathematical account of
mutual modeling but to be more systematic in describing the following
experiments.

@@

We studied the impact of an awareness tool on group performance and mutual
modeling~\citep{nova2007collaboration}. The availability
of an awareness tool was our independent variable. In previous studies, we
replayed a video of the game to subjects who surprised us by their ability to
remember former states of their mutual model: ``I did that because I thought that
you would do that''. Hence, this experiment focused the representation of each
other's action plans. During the game, we asked them to anticipate the next
action of their partner as well as to announce their own action. It is clear
asking them a question about their partner increases their attention for the
rest of the game.

@@

Task performance was measured by the score reached by the two subjects after
three situations. The effort of mutual modeling was measured as the ratio of
time that players would spend in the scout mode (divided by total time), which
is the time during which players are not performing their own actions but
monitoring their partner's actions.

@@

Ninety students participated in this experiment. We chosen only students from
our university because knowing the campus geography had an impact both on group
performance but also on mutual modeling: it is hard to represent the path of
someone across some space without having a mental map of this space. We selected
groups of students who knew each other. We assigned 10 triads to each of our
three experimental conditions. Each condition was made up of approximately 25\%
of women, but we did not control gender repartition within each triad.

@@

\gModel{A}{B} is the sum of errors made by A about B's paths. An error was
either drawing a place where the partner had not been or not drawing a place
where he/she had gone. One could argue that \gModel{A}{B} is biased by the
subjects' ability to translate their trajectories souvenir into a map drawing.
However, 85\% of subjects made no mistake at all when drawing their own path. We
therefore consider mistakes in their partners' path as being due to a lack of
mutual modeling accuracy instead of being due to spatial reasoning skills.

@@

This positive result regarding to the symmetry question is even more surprising
than the previous one: despite the high heterogeneity of spatial skills among
adults~\citep{liben1981spatial}, a significant correlation could indicate that
mutual modeling relies more on group processes than on individual processes. The
significant results on the triangle questions probably go in the same direction:
either the modeler or the modeled both contribute to accurate modeling, or that
the quality of teamwork has a global effect. In this experiment, since team
members did not interact intensively during the task, the factor that
facilitates accurate mutual modeling is probably less the quality of verbal
interactions that the group strategy: a clear strategy facilitates memorizing
one's partner path.

@@

\paragraph{Symmetry question} Regarding the main items, the mutual modeling of
emotions, the correlation is .25 or .36  \fixme{a reprendre avec - Patrick}

@@

\paragraph{Rectangle question}$\Delta_4$: the correlation between \gmodel{A}{A}
and \gmodel{A}{B} is globally not significant ($r=0.05$) as well as not
significant in each condition: someone good at self-modeling is not necessarily
good at modeling someone else and vice-versa.

@@

Despite the fact that the learning task was the same as in study 4, the
conditions of collaboration (viewing multiple maps or not) and the conditions
(scripted or not, awareness tool or not) probably explain differences in terms
of mutual modeling.

@@

Out of 5 studies, 3 revealed a correlation between the model peers built about
each other: when A builds an accurate model of B, B also tends to build an
accurate model of A.  In one study, this was the case only in one condition.
The conclusion is hence that modeling is mutual but only in certain conditions,
summarized in Table~\ref{synthesis_table}.

@@

Coming back to our initial hypotheses, the $\Delta_2$ results in studies 1, 2
and 3 reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The
\emph{interaction hypothesis} $\mathcal{H}_{3}$ was that the activity of
modeling one's partner is not simply the result of an individual cognitive
effort but overall the side-effect of the quality of interactions. The
\emph{second-order-modeling hypothesis} $\mathcal{H}_{2}$ stated that, if A is
accurate in modeling B, A should notice when B misunderstands A and should be
able to repair. A good modeler would somehow drive the modeling of his or her
partner.  This \emph{drive} will however occur through verbal cues and repairs
and is hence not easy to differentiate from $\mathcal{H}_{3}$.  The
contradictory $\Delta_1$ results in studies 4 (to some extent) and 5 weaken the
hypothesis.

The triangular questions lead to similar contradictions. The $\Delta_2$
correlation between the \gmodel{A}{C} and \gmodel{B}{C} (studies 2 and 3?)
support $\mathcal{H}_{2}$ if modeling skills were purely individual, there would
be no reason for modeling skills of A and B to be correlated.  The role of C in
\gmodel{A}{C} and \gmodel{B}{C} probably explains this correlation.  Conversely,
the $\Delta_3$ correlation between the \gmodel{A}{B} and \gmodel{A}{C} (studies
3 and 2?) question $\mathcal{H}_{3}$: if the role of the modellee was so
important, it would lead be surprising to have by chance a correlation between
B's and C' ability to drive A's modeling.

In addition, the lack of correlation between \gmodel{A}{A} and \gmodel{A}{B}
($\Delta_2$ in study 4 and study 5) could also contributes to the idea that A's
modeling skills is not the key factor. However, between modeling someone and
modeling oneself, \ie  metacognition, they are so many differences, that we
would hardly conclude from that,

@@

These conclusions must be presented with multiple disclaimers. First, they
heavily rely on correlations; hence we cannot identify causal links. Second, we
faced hard methodological issues. Providing learners with on-task questionnaires
introduce a bias: they will pay more attention to their partners in the
remaining time. Providing them with ``after-task'' questionnaires implies mnemonic
and rationalization biases. The nature of mutual modeling implies methodological
challenges that call for new  measurement methods. We have promising results for
using eye tracking methods to address this challenge.

Finally, let us stress the expressions we used, such as \gmodel{A}{B}, do not
imply we have a mechanical view of modeling. It's just a language for referring
to different components of modeling each other in teams.  This simple notation
emerges from the need to compare five mutual modelling studies; it did not
generate a consistent program of consistent studies.  Our results are not very
robust because they emerge from a \emph{post-hoc} reinterpretation of studies
that addressed different research questions (media richness, awareness tools,
scripts,...). This diversity makes our results difficult to integrate, partly
contradictory. Nonetheless, this diversity gives some generalizability: mutual
modelling has been investigated in different contexts (virtual space versus real
space), with different groups sizes (pairs and triads) and different tasks.
Moreover, these results have been found using different methods for measuring
modeling: on-task in study 1 versus off-task in study 2, subjective validation
in study 1 versus objective validation in study 2 (comparing A's model with B's
behavior).

>>>>

\title{Is Partner Modelling Mutual?}

@@

Deep grounding or shared meaning making requires some cognitive load. For
\citet{clark1986referring}, what is important is not the individual effort made
by the receiver of a communicative act, but the overall \emph{least collaborative
effort}.  The cost of producing a perfect utterance may be higher than the cost
of repairing the problems that may arise through misunderstandings. For
instance, subjects are less careful about adapting their utterances to their
partner when they know they can provide feedback on his/her
understanding~\citep{schober1993spatial}. We introduced the notion of
\emph{optimal collaborative effort}~\citep{dillenbourg1995evolution} to stress
that misunderstanding should not be viewed as something to be avoided (if this
was possible), but as an opportunity to engage into verbalization, explanation,
negotiation, and so forth. This issue is related to the global argument
regarding cognitive load in learning activities, especially in discovery
learning environments: there is no learning without some cognitive load, but
overload may hinder learning~\citep{paas2003cognitive}. In the context of
collaborative learning, we understand the cognitive load induced by mutual
modelling as part of~\citet{schwartz1995emergence} notion of effort towards a
shared understanding. For instance, CSCL researchers expanded the use of
\emph{collaboration scripts} \fixme{REF (Aronson,?)}. A script is a pedagogical
method that collaborative learning activities in order to foster the emergence
of productive interactions such as argumentation, explanation or conflict.
Conflict-resolution scripts such as the {\sc
ArgueGraph}~\citep{dillenbourg2008mechanics} forms pair of students with
opposite opinions, which increases the difficulty of consensus building,
requiring more justifications, more negotiation, more load. Similarly, JIGSAW
scripts provide peers with different but complementary knowledge for augmenting
(reasonably) the effort group members have to engage to reach a shared solution.

@@

\section{Mutual Modelling}

In order to build a shared understanding, do partners have to build a
representation of each other's knowledge? We refer to \emph{mutual modelling} as
the process of inferring one's partner mental states. Any claim that students
carry out a detailed monitoring of their peers would be as incorrect as any
claim that they do not maintain any representation at all. If mutual modelling
had to be permanently detailed and accurate, subjects would obviously face a
huge cognitive load. Conversely, peers could not collaborate without some
minimal amount of mutual modelling. For instance, A cannot disagree with B
without knowing that B has a different opinion. The mutual model can be implicit
(A is not aware of what he knows about B), by-default (I believe that B beliefs
what I believe unless contrary evidence), opportunistic (A does not model B
unless the conversation requires it), global (A infers B's beliefs based on
categories such as age, culture or profession) and, of course, it can be
incorrect… but it can not remain empty. Dialogues include many instances of
utterances such as ``I thought he would do that'' (first level of mutual
modelling) or even ``He thought I would do that but I intended something else.''
(second level of mutual modelling).

@@

A partner model is probably not a ``box'', \ie not a monolithical representation
but rather a mosaic of information fragments about the partner, with various
grain sizes and various life cycles. This mosaic is elaborated through various
mechanisms, first for building an initial model of the partner and then for
updating this model.  As two students meet for the first time, mutual models are
initialized by the assumptions they make upon each other from cues such as
his/her membership to large categories (age, culture, profession, ...) include
stereotypes (sportsmen, junkie, business women, Swiss,...) as well as physical
appearance. Scholars studied how initial modelling impacts communication. In
their experiments on initial mutual modelling, Slugoski, Lalljee, Lamb \&
Ginsburg~\citep{slugoski1993attribution} pretended to their subjects that their
(fake) partner had or had not received the same information. They observed that
the subjects adapted their dialogue by focusing the explanation on the items
that (s)he was supposed to ignore. Brennan (1991) \fixme{ref} showed that the subjects used
different initial strategies in forming queries depending on who they were told
their partner was.

@@

After initialization, mutual models are updated along the collaborative work
through verbal and non-verbal interactions. A default inference rule is that ``my
partner agrees with me unless he disagrees'', which reject the critiques that
mutual modelling generates an unbearable cognitive load. This default rule is
superseded by the several mechanisms for monitoring and repairing one's partner
understanding: acknowledgement, continuous attention, relevance of next turns,
facial expressions including gaze signals, etc.

Mutual modelling does not occur in a vacuum but it is highly contextualized. REF
reviews the features of the collaborative situation, namely the media
(cotemporality… ), may facilitate or hamper mutual modelling. Hutchins (SILENCE)
reported a study in which a short silence between two pilots was perfectly
interpreted because it occurred in a highly constrained communication context.
Some environments are more productive than others in helping peers to detect
their misunderstandings. Roschelle~\citep{roschelle1995construction} reformulate
the design of CSCL interfaces as providing ways for peer to detect and repairs
their misunderstanding. In CSCW, researchers have explored various so-called
``awareness tools'', \ie functionalities that inform A is about B's actions that A
could not directly perceived because B was working on a different subset of the
virtual space. Different awareness tools will be addressed in this contribution
since they allowed us to manipulate experimentally the mutual modelling activity.

@@

To report experiments on mutual modelling, we suggest
the notation \model{A}{B}{X} to denote ``A knows that B knows X''. This is a
simple but useful reduction to reason on mutual modelling. This
notation does not mean A has an explicit, monolithic representation of B: it
must be understood as an abstraction referring to complex socio-cognitive
processes. Besides, we refer to the \emph{degree of accuracy} of the model as
\Model{A}{B}{X}.

@@

    \item $cost(\alpha)$ In all studies but study 4, the cost of  grounding is
        an independent variable. Study 3 uses media richness as independent
        variable, with the hypothesis that modelling emotions is ``cheaper'' with a
        richer medium, \ie  when peers can see each other.  Studies 1,2 and 4
        use awareness tools which, in principle, reduce the cost of mutual modelling, but do
        not eliminate all costs: if the tool provide A with information about
        what B does/knows, this additional information may actually increase
        cognitive load. Awareness tools constitute as a kind of mutual modelling prosthesis,
        and, like any prosthesis, they many augment mutual modelling (by facilitating it or
        even scaffolding it) or inhibit it (by making it useless).

@@

\begin{itemize}
    \item {\bf Capturing \model{A}{B}{X}}: The simplest method is to ask $A$ what (s)he
        believes about what $B$ knows, feels, intends to do, etc. This raises
        obvious methodological concerns since such a question triggers a
        modelling process beyond what it would 'naturally' be. To avoid this
        bias, one can estimate mutual modelling after task completion. Then, the
        obvious drawback are memory losses and by post-hoc reconstruction.  The
        first method was used in study 2 and the second one in the other
        studies. Another option would be to rely on external behavioural metrics
        like eye-tracking: we hypothesise for instance that the fixation time
        reflects the efforts engaged by the human to understand, hence, model,
        the others. Such an approach has however not yet been investigated in
        the presented studies.

@@

$\mathcal{H}_{2}$ is also called a second level modelling hypothesis since H
needs to monitor B to see if A understood him/her
$\mathcal{M}(B,(A,\mathcal{M}(A,B)))$

Our main issue is the symmetry questions: what is the relationship
(Fig.~\ref{mm_symmetry}) between \Model{A}{B}{X} and \Model{B}{A}{X}? A low
symmetry would mean that mutual modelling is mainly an individual
attitude/aptitude ($\mathcal{H}_{1}$). A high correlation might support
$\mathcal{H}_{2}$ since there is a low probability that randomly formed pairs
integrate peers with the same level of mutual modelling skills.

@@

This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do A and
B have the same accuracy when modelling C? $\Delta_2 = \Delta(\M{A}{C}{X},
\M{B}{C}{X})$. If it is the same, $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ gain
over $\mathcal{H}_{1}$

@@

We could go further by comparing self- versus other modelling ($\Delta_4$ in
Fig.~\ref{mm_rectangle}) as an indication of meta-cognitive skills. We can also
question if modelling skills depend upon what aspects are being modeled ($X$ or $Y$),
which would explain vertical differences ($\Delta_5$ in
Fig.~\ref{mm_rectangle}). We do not however further develop these questions because we
don't have the necessary data in our studies.

@@

This simple notation does not pretend to provide a mathematical account of
mutual modelling but to be more systematic in describing the following
experiments.

@@

We studied the impact of an awareness tool on group performance and mutual
modelling~\citep{nova2007collaboration}. The availability
of an awareness tool was our independent variable. In previous studies, we
replayed a video of the game to subjects who surprised us by their ability to
remember former states of their mutual model: ``I did that because I thought that
you would do that''. Hence, this experiment focused the representation of each
other's action plans. During the game, we asked them to anticipate the next
action of their partner as well as to announce their own action. It is clear
asking them a question about their partner increases their attention for the
rest of the game.

@@

Task performance was measured by the score reached by the two subjects after
three situations. The effort of mutual modelling was measured as the ratio of
time that players would spend in the scout mode (divided by total time), which
is the time during which players are not performing their own actions but
monitoring their partner's actions.

@@

Ninety students participated in this experiment. We chosen only students from
our university because knowing the campus geography had an impact both on group
performance but also on mutual modelling: it is hard to represent the path of
someone across some space without having a mental map of this space. We selected
groups of students who knew each other. We assigned 10 triads to each of our
three experimental conditions. Each condition was made up of approximately 25\%
of women, but we did not control gender repartition within each triad.

@@

\gModel{A}{B} is the sum of errors made by A about B's paths. An error was
either drawing a place where the partner had not been or not drawing a place
where he/she had gone. One could argue that \gModel{A}{B} is biased by the
subjects' ability to translate their trajectories souvenir into a map drawing.
However, 85\% of subjects made no mistake at all when drawing their own path. We
therefore consider mistakes in their partners' path as being due to a lack of
mutual modelling accuracy instead of being due to spatial reasoning skills.

@@

This positive result regarding to the symmetry question is even more surprising
than the previous one: despite the high heterogeneity of spatial skills among
adults~\citep{liben1981spatial}, a significant correlation could indicate that
mutual modelling relies more on group processes than on individual processes. The
significant results on the triangle questions probably go in the same direction:
either the modeler or the modeled both contribute to accurate modelling, or that
the quality of teamwork has a global effect. In this experiment, since team
members did not interact intensively during the task, the factor that
facilitates accurate mutual modelling is probably less the quality of verbal
interactions that the group strategy: a clear strategy facilitates memorizing
one's partner path.

@@

\paragraph{Symmetry question} Regarding the main items, the mutual modelling of
emotions, the correlation is .25 or .36  \fixme{a reprendre avec - Patrick}

@@

\paragraph{Rectangle question}$\Delta_4$: the correlation between \gmodel{A}{A}
and \gmodel{A}{B} is globally not significant ($r=0.05$) as well as not
significant in each condition: someone good at self-modelling is not necessarily
good at modelling someone else and vice-versa.

@@

Despite the fact that the learning task was the same as in study 4, the
conditions of collaboration (viewing multiple maps or not) and the conditions
(scripted or not, awareness tool or not) probably explain differences in terms
of mutual modelling.

@@

Out of 5 studies, 3 revealed a correlation between the model peers built about
each other: when A builds an accurate model of B, B also tends to build an
accurate model of A.  In one study, this was the case only in one condition.
The conclusion is hence that modelling is mutual but only in certain conditions,
summarized in Table~\ref{synthesis_table}.

@@

Coming back to our initial hypotheses, the $\Delta_2$ results in studies 1, 2
and 3 reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The
\emph{interaction hypothesis} $\mathcal{H}_{3}$ was that the activity of
modelling one's partner is not simply the result of an individual cognitive
effort but overall the side-effect of the quality of interactions. The
\emph{second-order-modelling hypothesis} $\mathcal{H}_{2}$ stated that, if A is
accurate in modelling B, A should notice when B misunderstands A and should be
able to repair. A good modeler would somehow drive the modelling of his or her
partner.  This \emph{drive} will however occur through verbal cues and repairs
and is hence not easy to differentiate from $\mathcal{H}_{3}$.  The
contradictory $\Delta_1$ results in studies 4 (to some extent) and 5 weaken the
hypothesis.

The triangular questions lead to similar contradictions. The $\Delta_2$
correlation between the \gmodel{A}{C} and \gmodel{B}{C} (studies 2 and 3?)
support $\mathcal{H}_{2}$ if modelling skills were purely individual, there would
be no reason for modelling skills of A and B to be correlated.  The role of C in
\gmodel{A}{C} and \gmodel{B}{C} probably explains this correlation.  Conversely,
the $\Delta_3$ correlation between the \gmodel{A}{B} and \gmodel{A}{C} (studies
3 and 2?) question $\mathcal{H}_{3}$: if the role of the modellee was so
important, it would lead be surprising to have by chance a correlation between
B's and C' ability to drive A's modelling.

In addition, the lack of correlation between \gmodel{A}{A} and \gmodel{A}{B}
($\Delta_2$ in study 4 and study 5) could also contributes to the idea that A's
modelling skills is not the key factor. However, between modelling someone and
modelling oneself, \ie  metacognition, they are so many differences, that we
would hardly conclude from that,

@@

These conclusions must be presented with multiple disclaimers. First, they
heavily rely on correlations; hence we cannot identify causal links. Second, we
faced hard methodological issues. Providing learners with on-task questionnaires
introduce a bias: they will pay more attention to their partners in the
remaining time. Providing them with ``after-task'' questionnaires implies mnemonic
and rationalization biases. The nature of mutual modelling implies methodological
challenges that call for new  measurement methods. We have promising results for
using eye tracking methods to address this challenge.

Finally, let us stress the expressions we used, such as \gmodel{A}{B}, do not
imply we have a mechanical view of modelling. It's just a language for referring
to different components of modelling each other in teams.  This simple notation
emerges from the need to compare five mutual modelling studies; it did not
generate a consistent program of consistent studies.  Our results are not very
robust because they emerge from a \emph{post-hoc} reinterpretation of studies
that addressed different research questions (media richness, awareness tools,
scripts,...). This diversity makes our results difficult to integrate, partly
contradictory. Nonetheless, this diversity gives some generalizability: mutual
modelling has been investigated in different contexts (virtual space versus real
space), with different groups sizes (pairs and triads) and different tasks.
Moreover, these results have been found using different methods for measuring
modelling: on-task in study 1 versus off-task in study 2, subjective validation
in study 1 versus objective validation in study 2 (comparing A's model with B's
behavior).
